---
title: "LeBar_Predictathon"
author: "Luke LeBar"
date: "`r Sys.Date()`"
output: html_document
---

# Libraries

```{r}
suppressPackageStartupMessages(library(BrAPI)) # T3 connection)
suppressPackageStartupMessages(library(purrr)) # handle lists
suppressPackageStartupMessages(library(tidyverse)) # data handling
suppressPackageStartupMessages(library(here)) # root directory
suppressPackageStartupMessages(library(apercu)) # look at lists
suppressPackageStartupMessages(library(progress)) # progress bar
```

# Set up root directory

TODO: need to move this script to analysis folder

```{r current dir}
here::i_am("data/phenotypic_data.Rmd")
source(here("code","useful_functions.R"))
```

# WheatCAP Predictathon: Phenotypic Data

## Accessing T3 Breedbase

```{r}
rm(list=ls())
library(BrAPI)
library(purrr)
library(tidyverse)

wheat <- createBrAPIConnection("wheat.triticeaetoolbox.org", is_breedbase=TRUE)

wheat$login("LukeLeBar", "Imagine2003")

#wheat$get("/germplasm")
```

## Accessing Trial Metadata

```{r}
focal_trials <- c("AWY1_DVPWA_2024", "TCAP_2025_MANKS", "25_Big6_SVREC_SVREC",
                  "OHRWW_2025_SPO", "CornellMaster_2025_McGowan", "24Crk_AY2-3",
                  "2025_AYT_Aurora","YT_Urb_25","STP1_2025_MCG")

check_val <- function(x) {
  if (is.null(x) || length(x) == 0) return(NA)
  return(x)
}
results_list <- list()
for(trial_name in focal_trials){
  resp <- wheat$get("/studies", query=list(studyName=trial_name))
  if (length(resp$data) == 0) {
    message(paste("Skipping: No data found for", trial_name))
    next
  }
  trial_metadata <- resp$data[[1]]
  row_data <- data.frame(
    trial_name    = trial_name,
    program       = check_val(trial_metadata$trialName),
    trial_id      = check_val(trial_metadata$studyDbId),
    location      = check_val(trial_metadata$locationName),
    year         = check_val(unlist(trial_metadata$seasons)),
    planting_date = check_val(trial_metadata$startDate),
    harvest_date  = check_val(trial_metadata$endDate),
    design        = check_val(trial_metadata$experimentalDesign$PUI),
    stringsAsFactors = FALSE
  )
  results_list[[trial_name]] <- row_data
}
focal_metadata <- do.call(rbind, results_list)
```

## CV00: Gather Program Phenotypic Data

### Grabbing all Trials for our Programs of Interest

```{r}
programs <- focal_metadata$program
all_trials_list <- list()
for (prog in programs) {
  resp <- wheat$get("/studies", query = list(programName=prog, pageSize=1000))
  if (length(resp$data) == 0) next
  ## create a data frame summarizing trial id, name, location, year, program
  prog_trials <- lapply(resp$data, function(x) {
    data.frame(id = x$studyDbId, 
               name = x$studyName, 
               location = x$locationName,
               year=paste(unlist(x$seasons), collapse=", "), 
               program = prog)
  })
  prog_trials <- do.call(rbind, prog_trials)
  ## if there are too many trials, select those from the same year and location as focal trial
  if(nrow(prog_trials) > 20){
    focal_years <- focal_metadata %>% filter(program == prog) |> pull(year)
    focal_locations <- focal_metadata %>% filter(program == prog) %>% pull(location) %>% unique()
    prog_trials <- prog_trials %>% 
      filter(year %in% focal_years | location %in% focal_locations)
  }
  
  all_trials_list[[prog]] <- prog_trials
}
## look first five elements
ap(all_trials_list)
lapply(all_trials_list, nrow)
all_trials <- bind_rows(all_trials_list, .id="focal_trial")
all_trials$isFocal <- ifelse(all_trials$name %in% focal_trials, TRUE, FALSE)
table(all_trials$isFocal)
all(focal_trials %in% all_trials$name)

## export trial names to pull genotypic data from it
p2f <- here("data","allTrials_metadata.csv") # path to file
write.table(all_trials, p2f, sep=",", row.names=F, quote=F)
```

### Yield Data for All Trials

```{r}
all_trials <- all_trials %>% filter(!name %in% focal_metadata$trial_name)
trial_ids <- all_trials$id
wheat <- getBrAPIConnection("T3/Wheat")
vars_rm <- "season"

all_observations <- data.frame()

pb <- txtProgressBar(min=1, max=length(trial_ids), style = 3)
## this trial id is long: 10703
## this trial id times out: 10710, 10876
trial_ids <- trial_ids[!trial_ids %in% c("10703","10710","10876")]

for(i in seq_along(trial_ids)){
  setTxtProgressBar(pb, i)
  #write(paste0("Processing trial ", i, " of ", length(trial_ids)), stdout())
  search <- try(wheat$search("/observations", body=list(studyDbIds = c(trial_ids[i]))))
  results <- search$combined_data
  results <- results %>%
    map(function(item){
      item[vars_rm] <- NULL
      return(item)
    }) %>% 
    bind_rows
  if(nrow(results) >0){
    results <- results %>% filter(observationVariableName == "Grain yield - kg/ha|CO_321:0001218")
    all_observations <- bind_rows(all_observations, results)
  }
}
close(pb)

## export all trials yield
p2f <- here("data","phenotypic","allTrials_yield_raw.csv") # path to file
write.table(all_observations, p2f, sep=",", row.names=F, quote=F)
```

Modifications compared to Luke's version:

-   Print the iteration number to track progress and see if some trials are causing problems
-   Test if trial data has yield variable
-   Replace `rbind` with `bind_rows` for handling when columns names differ

## CV0: Pull Phenotype Data for Focal Accessions

```{r}
wheat <- getBrAPIConnection("T3/WheatCAP")
focal_acc <- wheat$wizard("accessions", list(trials=c(focal_trials)), verbose=T)
focal_acc <- as.data.frame(cbind(focal_acc$data$ids, focal_acc$data$names))
colnames(focal_acc)[1] = "acc_id"
colnames(focal_acc)[2] = "acc_name"
focal_ids <- focal_acc$acc_id
wheat <- getBrAPIConnection("T3/Wheat")
focal_pheno <- wheat$search("/observations", body=list(germplasmDbId=c(focal_ids)))

results <- focal_pheno$combined_data
results <- results %>%
  map(function(item){
  item[vars_rm] <- NULL
  return(item)
  }) %>% 
  bind_rows
results <- results %>% filter(observationVariableName == "Grain yield - kg/ha|CO_321:0001218")

write.csv(results, "focal_yield.csv")
```
