---
title: "Apply genomic prediction"
author: Charlotte Brault
output: html_document
df_print: kable
date: '`r format(Sys.time(), "%d %B, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      engine.opts = list(bash = "-l"))
```

# Libraries

```{r load libraries}
suppressPackageStartupMessages(library(here)) # root directory
suppressPackageStartupMessages(library(data.table)) # load data
suppressPackageStartupMessages(library(apercu)) # ap
suppressPackageStartupMessages(library(rrBLUP)) # genomic selection
suppressPackageStartupMessages(library(dplyr)) # formatting
suppressPackageStartupMessages(library(ggplot2)) # plot
suppressPackageStartupMessages(library(paletteer)) # plot palette
suppressPackageStartupMessages(library(janitor)) # formating
suppressPackageStartupMessages(library(doParallel)) # for parallelization
suppressPackageStartupMessages(library(foreach)) # for parallelization
suppressPackageStartupMessages(library(snow)) # for creating cluster
suppressPackageStartupMessages(library(doSNOW)) # for creating cluster
suppressPackageStartupMessages(library(purrr)) # list handling
suppressPackageStartupMessages(library(bigstatsr)) # ggplot theme
suppressPackageStartupMessages(library(ggbeeswarm)) # ggplot geom_quasirandom
#devtools::install_github("chabrault/HRSWGS")
#suppressPackageStartupMessages(library(HRSWGS)) ## custom functions

nb.cores <- 6
```


# Set up root directory

```{r current dir}
here::i_am("analysis/genom_pred.Rmd")
source(here("code","useful_functions.R")) # getFolds function
```


This R chunk is used to assess how much time it takes to execute the R code in this document until the end:

```{r time_0}
t0 <- proc.time()
```

```{r}
target.pred <- c("AWY1_DVPWA_2024","TCAP_2025_MANKS","25_Big6_SVREC_SVREC",
                 "OHRWW_2025_SPO","CornellMaster_2025_McGowan","24Crk_AY2-3",
                 "2025_AYT_Aurora","YT_Urb_25","STP1_2025_MCG")
tri <- "2025_AYT_Aurora"
```

# Load data

## Genomic data

```{r load_geno}
# p2fs <- here("data","genotypic", trials_predictathon,"clean",
#              paste0(trials_predictathon,"_focalAccessions_GRM.rds"))
p2fs <- here("data","genotypic", target.pred,"clean",
             paste0(target.pred,"_combined_GRM.rds"))
#              paste0(trials_predictathon,"_focalAccessions_GRM.rds"))
p2fs <- here("data","genotypic", target.pred,"clean",
             paste0(target.pred,"_Predictathon_genotypes_curated_numeric.rds"))

file.mtime(p2fs)
## for missing ones, use the focal accessions GRM
# missing <- which(is.na(file.exists(p2fs)))
# if(length(missing)>0){
#   p2fs[missing] <- here("data","genotypic", target.pred[missing],"clean",
#                         paste0(target.pred[missing],"_focalAccessions_GRM.rds"))
# }

genoList <- lapply(p2fs, readRDS)
names(genoList) <- target.pred
ap(genoList)
sapply(genoList, dim)


grmList <- lapply(genoList, function(x) x=rrBLUP::A.mat(x-1))
ap(grmList)

```

Small number of genotypes for "AWY1_DVPWA_2024"


## Phenotypic data



```{r load_pheno_trials}
p2f <- here("data","phenotypic","allAccessions_yield_raw.csv")
file.mtime(p2f)
pheno_init <- fread(p2f)
dim(pheno_init)
head(pheno_init)
str(pheno_init)
pheno <- pheno_init |> 
  rename(GID=germplasmName, YLD=value, PROG=focal_trial, LOC=location, YEAR=year) |> 
  mutate(ENV=paste0(LOC,"_",YEAR)) |>
  select(GID, PROG, LOC, YEAR, ENV, YLD)
```

## Common genotypes

```{r}
grmList.cmon <- purrr::map(grmList, \(x){
  int <- intersect(rownames(x), pheno$GID)
  return(x[int,int])
})
lapply(grmList.cmon, nrow)
```


# GBLUP CV0

cross validation by program.
Predict unknown trial with known genotypes. 
Steps:

- Subset the lines both genotyped and phenotyped for each target trial
- Predict each environment in turn using the other trials as training set.

```{r}
tri="TCAP_2025_MANKS"

res.cv0 <- expand.grid(trial=target.pred,  corP=NA,
                       nb.genos=NA, nb.envs=NA)


for(tri in target.pred){
  write(paste0("Processing trial: ", tri,"\n"), stdout())
  
  ## subset GID with geno and pheno data
  gid.geno <- rownames(grmList[[tri]])
  gid <- intersect(gid.geno, pheno$GID)
  length(gid)
  if(length(gid) == 0){
    write(paste0("No genotypes with both geno and pheno data for trial ", tri), stdout())
    
  }else {
    ## subset pheno and geno from the program to predict
    geno_sub <- grmList[[tri]][gid,gid]
    pheno_sub <- pheno |> filter(GID %in% gid)
    ## select environments with more than 50 obs
    envs <- pheno_sub %>% 
      group_by(ENV) %>% summarize(n=n(),.groups="drop") %>% 
      filter(n > 50) %>% pull(ENV)
    pheno_sub <- pheno_sub %>% filter(ENV %in% envs)
    
    df_pred_e <- data.frame()
    ## for loop over environments to predict each environment in turn
    for(e in envs){
      ## set yield data for the test environment to NA
      pheno_train <- pheno_sub |> 
        mutate(YLD=ifelse(ENV %in% e, NA, YLD))
      ## fit GBLUP model
      ### add fixed effects of location and year instead of env?
      fit <- rrBLUP::kin.blup(data=pheno_train, geno="GID",pheno="YLD", 
                              K=geno_sub, 
                              fixed=c("LOC","YEAR"),#"ENV", ## add env as fixed effect
                              n.core=nb.cores)
      ## export predicted values
      df_pred_e <- rbind(df_pred_e, 
                         data.frame(ENV=e, GID=names(fit$pred), pred=c(fit$pred)))
      
    }
    ## add observed values
    df_pred_e <- df_pred_e %>% left_join(pheno_sub, by=c("GID","ENV"))
    ## calculate correlation across environments
    pa <- cor(df_pred_e$YLD, df_pred_e$pred, use="complete.obs")
    idx <- which(res.cv0$trial == tri)
    res.cv0$corP[idx] <- pa
    res.cv0$nb.envs[idx] <- length(envs)
    
    ## add training/validation set size of the last fold
    res.cv0$nb.genos[idx] <- length(unique(pheno_train$GID))
    
    
  }
}

```


# GBLUP CV00

```{r}

res.cv00 <- expand.grid(trial=target.pred,  corP=NA,
                        nb.genos=NA, nb.envs=NA)
## subset GID with geno and pheno data
gid.geno <- rownames(grmList[[tri]])
gid <- intersect(gid.geno, pheno$GID)
length(gid)
# if(length(gid) == 0){
#   write(paste0("No genotypes with both geno and pheno data for trial ", tri), stdout())
#   
# }else {
## subset pheno and geno from the program to predict
geno_sub <- grmList[[tri]][gid,gid]
pheno_sub <- pheno |> filter(GID %in% gid)
df_pred_e <- data.frame()

## select environments with more than 50 obs
envs <- pheno_sub %>% 
  group_by(ENV) %>% summarize(n=n(),.groups="drop") %>% 
  filter(n > 50) %>% pull(ENV)
pheno_sub <- pheno_sub %>% filter(ENV %in% envs)


## for loop over environments to predict each environment in turn
for(e in envs){
  ## set yield data for the test environment to NA +
  ### set genotypes in the test environment as NA for all other environments
  gid.env <- pheno_sub$GID[pheno_sub$ENV %in% e]
  pheno_train <- pheno_sub |> 
    mutate(YLD=ifelse(ENV %in% e, NA, YLD),
           YLD=ifelse(GID %in% gid.env, NA, YLD))
  ## fit GBLUP model
  ### add fixed effects of location and year instead of env?
  fit <- rrBLUP::kin.blup(data=pheno_train, geno="GID",pheno="YLD", 
                          K=geno_sub, 
                          fixed=c("LOC","YEAR"),#"ENV", ## add env as fixed effect
                          n.core=nb.cores)
  ## export predicted values
  df_pred_e <- rbind(df_pred_e, 
                     data.frame(ENV=e, GID=names(fit$pred), pred=c(fit$pred)))
}
## add observed values
df_pred_e <- df_pred_e %>% left_join(pheno_sub, by=c("GID","ENV"))
## correlation within env
df_pred_e |> group_by(ENV) |>
  summarize(corP=cor(YLD, pred, use="complete.obs"),
            n=n_distinct(GID))
## calculate correlation across environments
pa <- cor(df_pred_e$YLD, df_pred_e$pred, use="complete.obs")
idx <- which(res.cv00$trial == tri)
res.cv00$corP[idx] <- pa
res.cv00$nb.envs[idx] <- length(envs)

## add training/validation set size of the last fold
res.cv00$nb.genos[idx] <- length(unique(pheno_train$GID))


#  }
```





<!-- ## Old  -->

<!-- ```{r} -->
<!-- tri=trials_predictathon[1] -->

<!-- res.cv0 <- expand.grid(trial=trials_predictathon, rep=1:nreps, corP=NA, -->
<!--                       nb.genos.TS=NA, nb.genos.VS=NA) -->


<!-- for(tri in trials_predictathon[3:length(trials_predictathon)]){ -->
<!--   write(paste0("Processing trial: ", tri,"\n"), stdout()) -->

<!--   ## subset GID with geno and pheno data -->
<!--   gid.geno <- rownames(grmList[[tri]]) -->
<!--   gid <- intersect(gid.geno, pheno$GID) -->
<!--   length(gid) -->
<!--   if(length(gid) == 0){ -->
<!--     write(paste0("No genotypes with both geno and pheno data for trial ", tri), stdout()) -->

<!--   }else { -->
<!--     geno_sub <- grmList[[tri]][gid,gid] -->
<!--     ## TODO subset phenotypes to the program to predict -->
<!--     pheno_sub <- pheno |> filter(GID %in% gid) -->
<!--     ## add mean pheno across environments for validation -->
<!--     pheno_avg <- pheno_sub %>% group_by(GID) %>%  -->
<!--       summarise(YLD=mean(YLD, na.rm=T)) -->

<!--     for(r in 1:nreps){ -->
<!--       write(paste0("  Rep: ", r,"\n"), stdout()) -->
<!--       ## determine testing set among gid genotypes -->
<!--       folds <- getFolds(index=length(gid), nb.folds=nfolds,seed=5656+r) -->
<!--       pred_f <- c() -->
<!--       for(f in 1:nfolds){ -->
<!--         test_gid <- gid[folds[[f]]] -->
<!--         train_gid <- setdiff(gid, test_gid) -->
<!--         pheno_train <- pheno_sub |> mutate(YLD=ifelse(GID %in% test_gid, NA, YLD)) -->
<!--         ## fit GBLUP model -->
<!--         fit <- rrBLUP::kin.blup(data=pheno_train, geno="GID",pheno="YLD",  -->
<!--                                 K=geno_sub, fixed="ENV", -->
<!--                                 n.core=nb.cores) -->
<!--         ## export predicted values -->
<!--         pred_f <- c(pred_f, fit$pred[test_gid]) -->

<!--       } -->
<!--       ## calculate correlation across folds -->
<!--       pa <- cor(pheno_avg$YLD, pred_f[match(pheno_avg$GID, names(pred_f))], use="complete.obs") -->
<!--       idx <- which(res.cv0$trial == tri & res.cv0$rep == r) -->
<!--       res.cv0$corP[idx] <- pa -->

<!--       ## add training/validation set size of the last fold -->
<!--       res.cv0$nb.genos.TS[idx] <- length(train_gid) -->
<!--       res.cv0$nb.genos.VS[idx] <- length(test_gid) -->

<!--     } -->
<!--   } -->
<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- tri <- trials_predictathon[2] -->
<!-- gid.geno <- rownames(grmList[[tri]]) -->
<!-- gid <- intersect(gid.geno, pheno$GID) -->
<!-- length(gid) -->
<!-- geno_sub <- grmList[[tri]][gid,gid] -->
<!-- pheno_sub <- pheno |> filter(GID %in% gid) -->
<!-- pheno_avg <- pheno_sub %>% group_by(GID) %>% summarise(YLD=mean(YLD, na.rm=T)) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- fit <- rrBLUP::kin.blup(data=pheno_sub, geno="GID",pheno="YLD",  -->
<!--                         K=geno_sub, fixed="ENV", -->
<!--                         n.core=nb.cores) -->
<!-- ap(fit$g) -->
<!-- ap(fit$pred) -->
<!-- cor(pheno_avg$YLD, fit$pred[match(pheno_avg$GID, names(fit$pred))], use="complete.obs") -->
<!-- ``` -->



<!-- # Genomic prediction -->

<!-- Cross-validation parameters: -->
<!-- ```{r CV_params} -->
<!-- dim(pheno) ## 1126 x 12 -->
<!-- stopifnot(!any(duplicated(pheno$GID))) -->
<!-- ## params: -->
<!-- ar <- "3k" ## genotypic array -->
<!-- nreps <- 10 ## number of cross-validation repetitions -->
<!-- nfolds <- 10 ## number of cross-validation folds -->
<!-- nIter <- 10000 ## default 6000 -->
<!-- burnIn <- 2000 ## default 1000 -->
<!-- h <- 1 ## bandwidth parameter for RKHS -->
<!-- rerun <- TRUE ## set to FALSE to avoid rerunning genomic prediction -->
<!-- ntraits <- ncol(pheno)-1 -->
<!-- (traits <- colnames(pheno)[-1]) ## verify that the trait names are correct -->
<!-- ## list of possible GP methods -->
<!-- methods <- c("rrBLUP","GBLUP","RKHS")#,"GBLUP",RKHS-KA","BayesA","BayesB","RandomForest","LASSO") -->
<!-- nmeths <- length(methods) -->
<!-- ``` -->

<!-- Apply genomic prediction for all methods and traits. -->

<!-- Description of the parameters: -->

<!-- * **geno**: genotype matrix, with individuals as rows and markers as columns. It is better to scale the genotype matrix. -->

<!-- * **pheno**: phenotype matrix, with individuals as rows and traits as columns, genotype name must be in a column named "GID". -->

<!-- * **traits**: vector of the names of the traits to predict. -->

<!-- * **GP.method**: method to use for genomic prediction. It can be one of the following: "rrBLUP", "RKHS", "RKHS-KA", "BayesA", "BayesB", "RandomForest", "LASSO". -->

<!-- * **nreps**: number of repetitions of the cross-validation. -->

<!-- * **nfolds**: number of folds of the cross-validation. -->

<!-- * **p2d.temp**: path to the directory where the temporary files will be saved for computing Bayesian methods. -->

<!-- * **h**: bandwidth for the RKHS method. It can be a single value or a vector of values to test. -->

<!-- * **p2f.stats**: path to the file where the predictive ability results will be saved. -->

<!-- * **nb.cores**: number of cores to use for parallelization. -->

<!-- ```{r, apply_gp_all} -->
<!-- gp.stats <- data.frame() -->
<!-- #pop <- paste0(pop_code.geno,"_",pop_code.trial,"_statgenSTA_",par.gv) -->
<!-- pop <- paste0(pop_code.trial,"_BLUPs") -->
<!-- for(meth in methods){ -->

<!--   p2f.stats <- here("output","genom_pred", -->
<!--                     paste0(pop,"_GP_",meth,"_results_",ntraits, "traits_intraCV.tsv")) -->
<!--   if(!file.exists(p2f.stats) |rerun){ -->
<!--     write(paste(meth,"\n"),stderr()) -->
<!--     out <- list() -->
<!--     out[[meth]] <-  compute_GP_methods(geno=geno, pheno=pheno, -->
<!--                                        traits=traits, GP.method=meth, -->
<!--                                        nreps=nreps, nfolds=nfolds,  -->
<!--                                        p2d.temp=here::here("output","genom_pred","temp"), -->
<!--                                        h=h,p2f.stats=p2f.stats,  -->
<!--                                        nb.cores=nb.cores) -->



<!--     #gp.statsT <- purrr::map_dfr(outT, function(x) x$gp.stats) -->
<!--     gp.stats <- rbind(gp.stats,out[[meth]]$gp.stats) -->

<!--   } else{ -->
<!--     gp.stats <- rbind(gp.stats, read.table(file=p2f.stats, sep="\t", header=TRUE)) -->

<!--   } -->
<!-- } -->

<!-- ap(gp.stats) -->
<!-- tapply(gp.stats$corP, gp.stats$trait, mean, na.rm=T) -->
<!-- ``` -->

<!-- Explore predicted vs. observed values -->

<!-- ```{r} -->
<!-- if(exists("out")){ -->
<!--   obspred <- do.call(rbind,out[[1]]$obspred$YIELD) -->

<!--   obspred <- obspred %>% group_by(GID) %>%  -->
<!--     summarise(obs=mean(obs), pred=mean(pred)) -->
<!--   #left_join(geno.info, by=GID) -->
<!--   plotly::ggplotly(ggplot(obspred, aes(x=obs, y=pred,label=GID))+ -->
<!--                      geom_point(alpha=0.6, size=1) +  -->
<!--                      geom_smooth(method="lm", se=FALSE, color="black")+ -->
<!--                      geom_rug(sides="bl")+ -->
<!--                      paletteer::scale_color_paletteer_d("ggthemes::Jewel_Bright") + -->
<!--                      ggpubr::stat_cor(method="pearson", size=3)+ -->
<!--                      geom_abline(intercept=0, slope=1, color="red")+ -->
<!--                      ggtitle("Predicted vs. observed values for MSU data", -->
<!--                              subtitle=paste0(methods[1]," genomic selection method"))+ -->
<!--                      theme_bigstatsr(size.rel=0.8)) -->
<!-- } -->

<!-- ``` -->



<!-- # Combine all  -->

<!-- ```{r} -->
<!-- all.res <- gp.stats -->
<!-- dim(all.res) -->
<!-- head(all.res) -->
<!-- all.res$GP.method <- factor(all.res$GP.method, levels=methods) -->
<!-- purrr::map_dbl(all.res, n_distinct) -->
<!-- tapply(all.res$corP, all.res$GP.method, mean) -->

<!-- tapply(all.res$corP, all.res$trait, mean) -->


<!-- summ.corP= all.res %>%  -->
<!--   group_by(trait) %>% summarise(meanCorP=mean(corP)) -->

<!-- print_table(summ.corP) -->
<!-- ntraits <- length(traits) -->
<!-- ##export the result table -->
<!-- p2f <- here("output","genom_pred", -->
<!--             paste0("GP_",nmeths,"meths_PredAb_",ntraits,"traits_intraCV.tsv")) -->
<!-- write.table(all.res, file=p2f, sep="\t", quote=FALSE, row.names=FALSE) -->
<!-- ``` -->


<!-- # Plot -->

<!-- ```{r} -->

<!-- # ggplot(all.res, aes(x=trait, y=corP, color=GP.method))+ -->
<!-- #   facet_wrap(~pop)+ -->
<!-- #   geom_boxplot() + theme_bw() -->
<!-- CPCOLS <- c("#008B00", "#EE82EE","red","lightseagreen", "#8B4500", "coral", "#4169E1") -->

<!-- ggplot(all.res, -->
<!--        aes(y=trait, x=corP, color=GP.method))+ -->
<!--   #geom_boxplot(position=position_dodge(0.9)) + -->
<!--   stat_summary(position=position_dodge(0.9),linewidth=1.2)+ -->
<!--   geom_point(alpha=0.3,position=position_dodge(0.9), size=1) + -->
<!--   #theme_bw() +  -->
<!--   theme_bigstatsr(size.rel=0.8)+ -->
<!--   xlab("Pearson correlation")+ -->
<!--   ggtitle("Predictive ability for MSU data", -->
<!--           subtitle=paste0(nmeths," methods 3K genotype array, 10-folds CV"))+ -->
<!--   scale_color_manual(values=CPCOLS, name="GP methods")+ -->
<!--   geom_hline(yintercept =seq(1.5,length(unique(all.res$trait))-0.5,1))+ -->
<!--   scale_x_continuous(breaks=seq(0,1,0.1), limits=c(0,1))+ -->
<!--   theme(axis.text.x=element_text(angle=0), -->
<!--         text=element_text(size=13), -->
<!--         panel.grid.major.y = element_blank(), -->
<!--         axis.ticks.y=element_blank(), -->
<!--         legend.position="bottom") -->
<!-- ggsave(file=here("output","figures", -->
<!--                  paste0("GP_PredAb_",nmeths,"meths_",ntraits,"traits_intraCV.png")), -->
<!--        height=8,width=11, scale=0.8) -->
<!-- ``` -->

<!-- # Predict genotypic values -->

<!-- Train the genomic selection model on the whole population using RKHS method and predict all the genotypic values. -->

<!-- ## Model training -->

<!-- ```{r pred_all_pop} -->
<!-- sel.meth <- "RKHS" ; h <- 1 ## bandwidth parameter for RKHS -->

<!-- p2d.temp=here::here("output","genom_pred","temp") -->
<!-- geno <- as.matrix(geno) -->
<!-- ## create the distance matrix (a bit similar to the GRM) for RKHS method -->
<!-- D <- as.matrix(dist(geno,method="euclidean"))^2 -->
<!-- D <- D/mean(D, na.rm=T) -->
<!-- ### compute kernel -->
<!-- K <- exp(-h*D) -->
<!-- dim(K) -->
<!-- ap(K) -->


<!-- all.pred <- data.frame() -->
<!-- for(tr in traits){ -->
<!--   write(tr,stderr()) -->

<!--   ## combine phenotypic data -->
<!--   # pheno <- rbind(phenoN[!phenoN$GID %in% cmon.pheno,c("GID",tr)], -->
<!--   #                phenoT[,c("GID",tr)]) -->
<!--   ## estimate marker effects on the whole population -->
<!--   yNA <- merge(data.frame(ind=rownames(geno)), -->
<!--                pheno[,c("GID",tr)], by.x="ind", by.y="GID", all.x=TRUE) -->
<!--   yNA <- yNA[[tr]] -->
<!--   p2f.temp <- paste0(p2d.temp,"/RKHS_all_",tr) -->

<!--   fit <- BGLR::BGLR(y=yNA,ETA=list(list(K=K,model='RKHS')), -->
<!--                     nIter=nIter,burnIn=burnIn,saveAt=p2f.temp,verbose=FALSE) -->
<!--   #as.data.frame(fit$yHat) -->
<!--   tmp <- data.frame(GID=rownames(geno), trait=tr, yHat=fit$yHat) -->
<!--   all.pred <- rbind(all.pred, tmp) -->
<!-- } -->

<!-- ## add genotype information -->
<!-- all.pred <- merge(all.pred, geno.info, by="GID",all.x=T, all.y=F) -->

<!-- ``` -->



<!-- ## Explore the prediction -->

<!-- Plot distribution -->
<!-- ```{r plot_distrib} -->
<!-- # all.pred$FIRST_YEAR[is.na(all.pred$FIRST_YEAR)] <- "Unknown" -->
<!-- # ggplot(all.pred, aes(x=FIRST_YEAR, y=yHat, color=FIRST_YEAR))+ -->
<!-- #   facet_wrap(~trait, scales="free_y")+ -->
<!-- #   geom_quasirandom(alpha=0.3,size=1, color="grey85") +  -->
<!-- #   geom_boxplot(fill=NA) + -->
<!-- #   paletteer::scale_color_paletteer_d("rcartocolor::Prism")+ -->
<!-- #   guides(color = guide_legend(override.aes = list(shape = 19)))+ -->
<!-- #   ggtitle("Genotypic value distribution", -->
<!-- #           subtitle="RKHS genomic selection method")+ -->
<!-- #   ylab("Predicted genotypic value")+xlab("First year")+ -->
<!-- #   theme_bigstatsr(size.rel=0.8)+  -->
<!-- #   guides(color=guide_legend(override.aes=list(shape=19,size=2),nrow=2))+ -->
<!-- #   scale_x_discrete(labels = scales::label_wrap(10))+ -->
<!-- #   theme(axis.text.x=element_blank(), -->
<!-- #         legend.position="inside", -->
<!-- #         legend.title=element_blank(), -->
<!-- #         legend.position.inside=c(0.65,0.12)) -->
<!-- #  -->
<!-- #  -->
<!-- # p2f <- here("output","figures", -->
<!-- #             paste0(pop,"_GP_PredVal_Distrib_",ntraits,"traits_wholeTrialPopGSTrain.png")) -->
<!-- # ggsave(file=p2f,height=10,width=13, scale=0.8) -->

<!-- ``` -->



<!-- Transform to a wide format -->

<!-- ```{r} -->
<!-- info.cols <- colnames(geno.info) -->

<!-- all.pred.wide <- all.pred %>%  -->
<!--   group_by(across(all_of(c(info.cols,"trait")))) %>% -->
<!--   summarize(yHat=mean(yHat, na.rm=T),.groups="drop") %>% -->
<!--   #select(all_of(c("geno","trait","pheno","Pedigree","Parent1","Parent2","EXPT","yHat")))%>% -->
<!--   tidyr::pivot_wider(names_from="trait", values_from="yHat") -->
<!-- head(all.pred.wide) -->



<!-- p2f <- here("output","genom_pred", -->
<!--             paste0(pop_code.geno,"_GP_PredVal_",sel.meth,"_", -->
<!--                    ntraits,"traits.tsv")) -->
<!-- write.table(all.pred.wide, file=p2f, row.names=F, sep="\t", col.names=T) -->


<!-- # Get the summary table for all traits -->
<!-- gtsummary::tbl_summary(all.pred.wide, include=all_of(traits), by=FIRST_YEAR, -->
<!--                        type = all_continuous() ~ "continuous2", -->
<!--                        statistic = all_continuous() ~ c( -->
<!--                          "{median} ({p25}, {p75})", -->
<!--                          "{min}, {max}" -->
<!--                        )) -->
<!-- #add_p(pvalue_fun = label_style_pvalue(digits = 2)) -->
<!-- ## print the highest 50 genotypic values for yield -->
<!-- all.pred.wide %>% arrange(desc(YIELD)) %>% head(200) %>% print_table() -->

<!-- ``` -->



<!-- ## Export observed and predicted values -->

<!-- ```{r export_obs_pred} -->
<!-- ## combine all phenotypic data -->
<!-- all.pheno <- pheno -->
<!-- dim(all.pheno) -->
<!-- obs.pred <- merge(all.pheno, all.pred.wide, by="GID", all=TRUE,suffixes=c(".Obs",".Pred")) -->
<!-- obs.pred <- relocate(obs.pred, all_of(info.cols)) -->
<!-- ## put observed and predicted values side by side -->
<!-- obs.pred <- reduce(.x = traits,.f = ~ relocate(.x, starts_with(.y)),.init = obs.pred) |>  -->
<!--   relocate(all_of(info.cols)) -->

<!-- ## put at first columns identifying the genotype -->
<!-- obs.pred <- relocate(obs.pred, GID) |> distinct() |>  -->
<!--   group_by(across(all_of(info.cols))) |> -->
<!--   summarize(across(where(is.numeric), ~ mean(.x, na.rm=T)),.groups="drop") |>  -->
<!--   arrange(desc(YIELD.Pred)) -->
<!-- ## print first 100 rows -->
<!-- print_table(head(obs.pred, n=100)) -->

<!-- ## remove some decimals to numeric columns -->
<!-- obs.pred <- obs.pred |>  -->
<!--   mutate(across(where(is.numeric), ~ round(.x, digits=2))) -->

<!-- p2f <- here("output", paste0(pop_code.geno,"_GP_ObsPredVal_",ntraits,"traits.tsv")) -->
<!-- write.table(obs.pred, file=p2f, row.names=F, sep="\t", col.names=T) -->
<!-- ``` -->


# Appendix

```{r info}
t1 <- proc.time(); t1 - t0
print(sessionInfo(), locale=FALSE)
```



