---
title: "Pull genotype data from T3"
author: Charlotte Brault/Marcos
output: html_document
date: '`r format(Sys.time(), "%d %B, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Libraries

```{r load libraries}
suppressPackageStartupMessages(library(here)) # root directory
suppressPackageStartupMessages(library(data.table)) # load data
suppressPackageStartupMessages(library(apercu)) # ap
suppressPackageStartupMessages(library(dplyr)) # formatting
suppressPackageStartupMessages(library(purrr)) # handle lists
## devtools::install_github("chabrault/HRSWGS") # custom package instal
#suppressPackageStartupMessages(library(HRSWGS)) # custom package
## devtools::install_github("TriticeaeToolbox/BrAPI.R")
suppressPackageStartupMessages(library(BrAPI)) # T3 connection
suppressPackageStartupMessages(library(T3BrapiHelpers)) # T3 helper
suppressPackageStartupMessages(library(DT)) # print table
suppressPackageStartupMessages(library(ggplot2)) # plot
suppressPackageStartupMessages(library(vcfR)) # load vcf
suppressPackageStartupMessages(library(progress)) # progress bar
suppressPackageStartupMessages(library(janitor))

```


# Set up root directory

```{r current dir}
here::i_am("analysis/pullGenoAssociatedTrials.Rmd")
source(here("code","useful_functions.R"))

here::here()
file.exists(here("code","useful_functions.R"))

```
# Create T3 connection

What's the difference between getBrAPIConnection and createBrAPIConnection?
```{r}
#wheat <- getBrAPIConnection("T3/Wheat")
db <- getBrAPIConnection("T3/Wheat")
wheat <- createBrAPIConnection("wheat.triticeaetoolbox.org", is_breedbase=TRUE)
#trials_df <- T3BrapiHelpers::getAllTrialMetaData(wheat,"Wheat")
# quick sanity checks
db$server
wheat$server

# test a tiny query (should return a data.frame)
tmp <- wheat$vcf_archived_list(genotyping_protocol_id = 291, verbose = FALSE)
str(tmp)

```

```{r}
trials_predictathon <- c("AWY1_DVPWA_2024","TCAP_2025_MANKS","25_Big6_SVREC_SVREC",
                         "OHRWW_2025_SPO","CornellMaster_2025_McGowan","24Crk_AY2-3",
                         "2025_AYT_Aurora","YT_Urb_25","STP1_2025_MCG")


# check that each focal trial folder exists and has manifest.csv
manifest_paths <- file.path(here("data","genotypic"), trials_predictathon, "manifest.csv")
data.frame(trial = trials_predictathon,
           manifest_exists = file.exists(manifest_paths),
           path = manifest_paths)

# actually load them
manifests <- purrr::map_dfr(trials_predictathon, \(x){
  data.table::fread(here("data","genotypic", x, "manifest.csv"))
})

manifests[, .(trial, protocol_id, protocol_name, project_id, project_name)]

```

#Setting up the dataset
```{r}

# --- read + clean metadata ---
all_trials_raw <- data.table::fread(here("data","allTrials_metadata.csv"))
all_trials <- all_trials_raw |>
  dplyr::select(-V1) |>
  dplyr::mutate(isFocal = as.logical(isFocal))

all_trials_focal    <- all_trials |> dplyr::filter(isFocal)
all_trials_nonfocal <- all_trials |> dplyr::filter(!isFocal)

```
#Helper clean name
```{r}
safe_filename <- function(x) {
  x <- basename(x)                      # in case the API returns a path
  x <- gsub("[<>:\"/\\\\|?*]", "_", x)  # Windows-illegal chars -> underscore
  x <- gsub("\\s+", "_", x)             # spaces -> underscore (optional)
  x
}

```


#Download WSU non-focal genotypic data
```{r}

prog <- "Washington State University"
wsu <- all_trials_nonfocal |> dplyr::filter(program == prog)

# holder columns
wsu <- wsu |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in 1:nrow(wsu)){
  tri_id <- wsu$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  wsu$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    wsu$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    wsu$best_protocol[i]    <- info$protocol_name
    wsu$best_protocol_id[i] <- info$protocol_id
    wsu$best_project[i]     <- info$project_name
    wsu$best_project_id[i]  <- info$project_id
    wsu$vcf_file_name[i]    <- info$file_name
  }
}

wsu_ready <- wsu |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(wsu)        # total WSU non-focal trials
nrow(wsu_ready)  # downloadable ones
wsu_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                           best_protocol_id, best_project_id, vcf_file_name) |> head(10)


###Download starts here 
prog <- "Washington State University"
focal_tri <- "AWY1_DVPWA_2024"  # WSU focal trial folder in your repo

# folder for this program under the focal trial
prog_slug <- "washington_state_university"
out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# download loop (skip if already exists)
wsu_dl <- wsu_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in seq_len(nrow(wsu_dl))){
  trial_id   <- wsu_dl$focal_trial[i]
  prot_id    <- wsu_dl$best_protocol_id[i]
  project_id <- wsu_dl$best_project_id[i]
  file_name  <- wsu_dl$vcf_file_name[i]

  file_name_clean <- safe_filename(file_name)

  out_file <- file.path(
    out_dir,
    paste0(
      "trial_", trial_id,
      "_prot_", prot_id,
      "_proj_", project_id,
      "_", file_name_clean
    )
  )

  wsu_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,   # keep original for API
      verbose = FALSE
    )
  }

  wsu_dl$downloaded[i] <- file.exists(out_file)
}

sum(wsu_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_wsu_nonfocal.csv")
write.csv(wsu_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)



```

#Download UMN non-focal genotypic data
```{r}

prog <- "University of Minnesota"
umn <- all_trials_nonfocal |> dplyr::filter(program == prog)

umn <- umn |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in 1:nrow(umn)){
  tri_id <- umn$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  umn$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    umn$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    umn$best_protocol[i]    <- info$protocol_name
    umn$best_protocol_id[i] <- info$protocol_id
    umn$best_project[i]     <- info$project_name
    umn$best_project_id[i]  <- info$project_id
    umn$vcf_file_name[i]    <- info$file_name
  }
}

umn_ready <- umn |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(umn)        # total UMN non-focal trials
nrow(umn_ready)  # downloadable ones
umn_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                           best_protocol_id, best_project_id, vcf_file_name) |> head(10)

###Download starts here 

prog <- "University of Minnesota"
focal_tri <- "24Crk_AY2-3"   # UMN focal trial folder

prog_slug <- "university_of_minnesota"
out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

umn_dl <- umn_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in 1:nrow(umn_dl)){
  trial_id   <- umn_dl$focal_trial[i]
  prot_id    <- umn_dl$best_protocol_id[i]
  project_id <- umn_dl$best_project_id[i]
  file_name  <- umn_dl$vcf_file_name[i]

  file_name_clean <- safe_filename(file_name)

  out_file <- file.path(
    out_dir,
    paste0(
      "trial_", trial_id,
      "_prot_", prot_id,
      "_proj_", project_id,
      "_", file_name_clean
    )
  )

  umn_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,     # IMPORTANT: keep original here for the API call
      verbose = FALSE
    )
  }

  umn_dl$downloaded[i] <- file.exists(out_file)
}


sum(umn_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_umn_nonfocal.csv")
write.csv(umn_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)



```
#Download SDSU non-focal genotypic data
```{r}
prog <- "South Dakota State University- Winter Wheat"
sdsu <- all_trials_nonfocal |> dplyr::filter(program == prog)

sdsu <- sdsu |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in 1:nrow(sdsu)){
  tri_id <- sdsu$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  sdsu$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    sdsu$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    sdsu$best_protocol[i]    <- info$protocol_name
    sdsu$best_protocol_id[i] <- info$protocol_id
    sdsu$best_project[i]     <- info$project_name
    sdsu$best_project_id[i]  <- info$project_id
    sdsu$vcf_file_name[i]    <- info$file_name
  }
}

sdsu_ready <- sdsu |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(sdsu)
nrow(sdsu_ready)
sdsu_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                            best_protocol_id, best_project_id, vcf_file_name)

#Download starts here 
focal_tri <- "2025_AYT_Aurora"
prog_slug <- "south_dakota_state_university_winter_wheat"

out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

sdsu_dl <- sdsu_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in seq_len(nrow(sdsu_dl))){
  trial_id   <- sdsu_dl$focal_trial[i]
  prot_id    <- sdsu_dl$best_protocol_id[i]
  project_id <- sdsu_dl$best_project_id[i]
  file_name  <- sdsu_dl$vcf_file_name[i]

  file_name_clean <- safe_filename(file_name)

  out_file <- file.path(
    out_dir,
    paste0(
      "trial_", trial_id,
      "_prot_", prot_id,
      "_proj_", project_id,
      "_", file_name_clean
    )
  )

  sdsu_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,   # keep original for API call
      verbose = FALSE
    )
  }

  sdsu_dl$downloaded[i] <- file.exists(out_file)
}

sum(sdsu_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_sdsu_nonfocal.csv")
write.csv(sdsu_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)


```
#Download TAMU non-focal genotypic data
```{r}
prog <- "TAMU"
tamu <- all_trials_nonfocal |> dplyr::filter(program == prog)

tamu <- tamu |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in 1:nrow(tamu)){
  tri_id <- tamu$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  tamu$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    tamu$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    tamu$best_protocol[i]    <- info$protocol_name
    tamu$best_protocol_id[i] <- info$protocol_id
    tamu$best_project[i]     <- info$project_name
    tamu$best_project_id[i]  <- info$project_id
    tamu$vcf_file_name[i]    <- info$file_name
  }
}

tamu_ready <- tamu |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(tamu)
nrow(tamu_ready)
tamu_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                            best_protocol_id, best_project_id, vcf_file_name)

#Download stats here
focal_tri <- "STP1_2025_MCG"
prog_slug <- "tamu"

out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

tamu_dl <- tamu_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in seq_len(nrow(tamu_dl))){
  trial_id   <- tamu_dl$focal_trial[i]
  prot_id    <- tamu_dl$best_protocol_id[i]
  project_id <- tamu_dl$best_project_id[i]
  file_name  <- tamu_dl$vcf_file_name[i]

  file_name_clean <- safe_filename(file_name)

  out_file <- file.path(
    out_dir,
    paste0(
      "trial_", trial_id,
      "_prot_", prot_id,
      "_proj_", project_id,
      "_", file_name_clean
    )
  )

  tamu_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,   # keep original for the API call
      verbose = FALSE
    )
  }

  tamu_dl$downloaded[i] <- file.exists(out_file)
}

sum(tamu_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_tamu_nonfocal.csv")
write.csv(tamu_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)



```

#Download Cornell non-focal genotypic data
```{r}
prog <- "Cornell University"
cornell <- all_trials_nonfocal |> dplyr::filter(program == prog)

cornell <- cornell |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in 1:nrow(cornell)){
  tri_id <- cornell$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  cornell$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    cornell$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    cornell$best_protocol[i]    <- info$protocol_name
    cornell$best_protocol_id[i] <- info$protocol_id
    cornell$best_project[i]     <- info$project_name
    cornell$best_project_id[i]  <- info$project_id
    cornell$vcf_file_name[i]    <- info$file_name
  }
}

cornell_ready <- cornell |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(cornell)
nrow(cornell_ready)
cornell_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                               best_protocol_id, best_project_id, vcf_file_name)

#Download starts here 

focal_tri <- "CornellMaster_2025_McGowan"
prog_slug <- "cornell_university"

out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

cornell_dl <- cornell_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in 1:nrow(cornell_dl)){
  trial_id   <- cornell_dl$focal_trial[i]
  prot_id    <- cornell_dl$best_protocol_id[i]
  project_id <- cornell_dl$best_project_id[i]
  file_name  <- cornell_dl$vcf_file_name[i]

  out_file <- file.path(
    out_dir,
    paste0("trial_", trial_id,
           "_prot_", prot_id,
           "_proj_", project_id,
           "_", file_name, ".vcf.gz")
  )

  cornell_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,
      verbose = FALSE
    )
  }

  cornell_dl$downloaded[i] <- file.exists(out_file)
}

sum(cornell_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_cornell_nonfocal.csv")
write.csv(cornell_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)



```
#Download Wisconsin non-focal genotypic data
```{r}
prog <- "University of Wisconsin"
uw <- all_trials_nonfocal |> dplyr::filter(program == prog)

uw <- uw |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in 1:nrow(uw)){
  tri_id <- uw$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  uw$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    uw$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    uw$best_protocol[i]    <- info$protocol_name
    uw$best_protocol_id[i] <- info$protocol_id
    uw$best_project[i]     <- info$project_name
    uw$best_project_id[i]  <- info$project_id
    uw$vcf_file_name[i]    <- info$file_name
  }
}

uw_ready <- uw |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(uw)
nrow(uw_ready)
uw_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                          best_protocol_id, best_project_id, vcf_file_name)

#Download starts here 

focal_tri <- "OHRWW_2025_SPO"
prog_slug <- "university_of_wisconsin"

out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

uw_dl <- uw_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in seq_len(nrow(uw_dl))){
  trial_id   <- uw_dl$focal_trial[i]
  prot_id    <- uw_dl$best_protocol_id[i]
  project_id <- uw_dl$best_project_id[i]
  file_name  <- uw_dl$vcf_file_name[i]

  file_name_clean <- safe_filename(file_name)

  out_file <- file.path(
    out_dir,
    paste0(
      "trial_", trial_id,
      "_prot_", prot_id,
      "_proj_", project_id,
      "_", file_name_clean
    )
  )

  uw_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,  # keep original for API call
      verbose = FALSE
    )
  }

  uw_dl$downloaded[i] <- file.exists(out_file)
}

sum(uw_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_uw_nonfocal.csv")
write.csv(uw_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)


```

#Download Michigan non-focal genotypic data
```{r}
prog <- "Michigan State University"
msu <- all_trials_nonfocal |> dplyr::filter(program == prog)

msu <- msu |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

for(i in seq_len(nrow(msu))){
  tri_id <- msu$focal_trial[i]

  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  msu$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    msu$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    msu$best_protocol[i]    <- info$protocol_name
    msu$best_protocol_id[i] <- info$protocol_id
    msu$best_project[i]     <- info$project_name
    msu$best_project_id[i]  <- info$project_id
    msu$vcf_file_name[i]    <- info$file_name
  }
}

msu_ready <- msu |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(msu)
nrow(msu_ready)
msu_ready |> dplyr::select(focal_trial, year, name, nb_genos, prop_geno,
                           best_protocol_id, best_project_id, vcf_file_name) |>
  dplyr::arrange(desc(year), name)

#Download starts here
focal_tri <- "25_Big6_SVREC_SVREC"
prog_slug <- "michigan_state_university"

out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

msu_dl <- msu_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

for(i in seq_len(nrow(msu_dl))){
  trial_id   <- msu_dl$focal_trial[i]
  prot_id    <- msu_dl$best_protocol_id[i]
  project_id <- msu_dl$best_project_id[i]
  file_name  <- msu_dl$vcf_file_name[i]

  file_name_clean <- safe_filename(file_name)

  out_file <- file.path(
    out_dir,
    paste0(
      "trial_", trial_id,
      "_prot_", prot_id,
      "_proj_", project_id,
      "_", file_name_clean
    )
  )

  msu_dl$out_file[i] <- out_file

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name = file_name,  # keep original for API call
      verbose = FALSE
    )
  }

  msu_dl$downloaded[i] <- file.exists(out_file)
}

sum(msu_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_msu_nonfocal.csv")
write.csv(msu_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)




```
#Download Illinois non-focal genotypic data
```{r}
prog <- "University of Illinois"
uiuc <- all_trials_nonfocal |> dplyr::filter(program == prog)

uiuc <- uiuc |>
  dplyr::mutate(
    nb_genos = NA_integer_,
    prop_geno = NA_real_,
    best_protocol = NA_character_,
    best_protocol_id = NA_character_,
    best_project = NA_character_,
    best_project_id = NA_character_,
    vcf_file_name = NA_character_
  )

pb <- txtProgressBar(min = 1, max = nrow(uiuc), style = 3)

for(i in 1:nrow(uiuc)){
  setTxtProgressBar(pb, i)

  tri_id <- uiuc$focal_trial[i]
  genos <- getGermplasmFromTrialVec(study_id_vec = tri_id, db)
  uiuc$nb_genos[i] <- nrow(genos)

  resp <- try(db$filter_geno_projects(genos$germplasmDbId), silent = TRUE)

  if(!inherits(resp, "try-error") && length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches[1]
    nb_genos  <- sum(resp$data$matches[, top_match], na.rm = TRUE)
    uiuc$prop_geno[i] <- 100 * (nb_genos / nrow(genos))

    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]

    info_all <- wheat$vcf_archived_list(genotyping_protocol_id = protId, verbose = FALSE) |> 
      as.data.frame()

    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    info <- info_all |> dplyr::filter(project_id == projId)
    if(nrow(info) == 0) info <- info_all[1, , drop = FALSE]
    if(nrow(info) > 1)  info <- info[1, , drop = FALSE]

    uiuc$best_protocol[i]    <- info$protocol_name
    uiuc$best_protocol_id[i] <- info$protocol_id
    uiuc$best_project[i]     <- info$project_name
    uiuc$best_project_id[i]  <- info$project_id
    uiuc$vcf_file_name[i]    <- info$file_name
  }
}
close(pb)

uiuc_ready <- uiuc |> dplyr::filter(!is.na(best_project_id) & !is.na(vcf_file_name))

nrow(uiuc)
nrow(uiuc_ready)

uiuc_ready |>
  dplyr::count(best_protocol_id, best_project_id, vcf_file_name, sort = TRUE) |>
  head(20)

#Download starts here
focal_tri <- "YT_Urb_25"
prog_slug <- "university_of_illinois"

out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

uiuc_dl <- uiuc_ready |> dplyr::mutate(out_file = NA_character_, downloaded = FALSE)

if(nrow(uiuc_dl) == 0){
  message("No downloadable UIUC non-focal trials found.")
} else {

  pb <- txtProgressBar(min = 1, max = nrow(uiuc_dl), style = 3)

  for(i in seq_len(nrow(uiuc_dl))){
    setTxtProgressBar(pb, i)

    trial_id   <- uiuc_dl$focal_trial[i]
    prot_id    <- uiuc_dl$best_protocol_id[i]
    project_id <- uiuc_dl$best_project_id[i]
    file_name  <- uiuc_dl$vcf_file_name[i]

    file_name_clean <- safe_filename(file_name)

    out_file <- file.path(
      out_dir,
      paste0(
        "trial_", trial_id,
        "_prot_", prot_id,
        "_proj_", project_id,
        "_", file_name_clean
      )
    )

    uiuc_dl$out_file[i] <- out_file

    if(!file.exists(out_file)){
      wheat$vcf_archived(
        output = out_file,
        genotyping_protocol_id = prot_id,
        genotyping_project_id  = project_id,
        file_name = file_name,  # keep original for API call
        verbose = FALSE
      )
    }

    uiuc_dl$downloaded[i] <- file.exists(out_file)
  }

  close(pb)
}

sum(uiuc_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_uiuc_nonfocal.csv")
write.csv(uiuc_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)


```

#Download Kansas non-focal genotypic data

```{r}

focal_tri <- "TCAP_2025_MANKS"
prog_slug <- "usda_ars_kansas"
protocol_ids <- c(289, 297)  # Allegro V1 and Allegro V2 (example)

# Output folder (same structure as your other programs)
out_dir <- here("data", "genotypic", focal_tri, "nonFocal", prog_slug, "raw_archived_vcfs")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# ---- 1) Build a download table (like *_ready / *_dl) ----
kansas_info <- lapply(protocol_ids, function(pid){
  wheat$vcf_archived_list(genotyping_protocol_id = pid, verbose = FALSE) |>
    as.data.frame() |>
    dplyr::mutate(protocol_id = pid)
}) |>
  dplyr::bind_rows()

stopifnot(nrow(kansas_info) > 0)

kansas_dl <- kansas_info |>
  dplyr::mutate(
    file_name_clean = vapply(file_name, safe_filename, FUN.VALUE = character(1)),
    out_file = file.path(
      out_dir,
      paste0(
        "prot_", protocol_id,
        "_proj_", project_id,
        "_", file_name_clean
      )
    ),
    downloaded = FALSE
  )

# ---- 2) Download loop (same idea as your others) ----
for(i in seq_len(nrow(kansas_dl))){
  prot_id    <- kansas_dl$protocol_id[i]
  project_id <- kansas_dl$project_id[i]
  file_name  <- kansas_dl$file_name[i]
  out_file   <- kansas_dl$out_file[i]

  if(!file.exists(out_file)){
    wheat$vcf_archived(
      output = out_file,
      genotyping_protocol_id = prot_id,
      genotyping_project_id  = project_id,
      file_name              = file_name,  # keep original for API call
      verbose = FALSE
    )
  }

  kansas_dl$downloaded[i] <- file.exists(out_file)
}

# Optional: file size
kansas_dl$file_size <- ifelse(kansas_dl$downloaded, file.info(kansas_dl$out_file)$size, NA_real_)

# ---- 3) Manifest (same pattern as your other programs) ----
sum(kansas_dl$downloaded)

manifest_path <- file.path(out_dir, "manifest_kansas_archived_vcfs.csv")
write.csv(kansas_dl, manifest_path, row.names = FALSE)

manifest_path
file.exists(manifest_path)
```


# Get genotyping platform used for each focal trial

```{r}
p2ds <- here("data","genotypic",trials_predictathon)
list.files(p2ds)
## load manifest files for each focal trial
manifests <- map_dfr(trials_predictathon, \(x){
  fread(here("data","genotypic",x,"manifest.csv"))
})
manifests
protid = unique(manifests$protocol_id)
```


# Get list of associated trials

```{r}
p2f <- here("data","allTrials_metadata.csv") # path to file
all_trials <- fread(p2f) |> 
  select(-V1) 
all_trials_focal <- all_trials |> 
  filter(isFocal) 
focal_prg <- unique(all_trials_focal$program)

all_trials <- all_trials |> filter(!isFocal)
head(all_trials)
```

# Pull genotyping platform for each associated trial

```{r}
all_trials$prop_geno <- NA 
all_trials$best_protocol <- NA
all_trials$best_protocol_id <- NA
all_trials$best_project <- NA
all_trials$best_project_id <- NA
all_trials$vcf_file_name <- NA

#tri=all_trials$focal_trial[165]
pb <- txtProgressBar(min=1, max=nrow(all_trials), style = 3)
for(i in 1:nrow(all_trials)){
  setTxtProgressBar(pb, i)
  tri <- all_trials$focal_trial[i]
  ## get the list of genotypes in this trial
  genos <- getGermplasmFromTrialVec(study_id_vec=tri, db)
  
  ## add the total number of lines
  all_trials$nb_genos[i] <- nrow(genos)
  ## which genotyping platform is the best?
  resp <- try(db$filter_geno_projects(genos$germplasmDbId))
  if(length(resp) == 4 && resp$status$category == "Success"){
    top_match <- resp$data$top_matches
    ## TODO if there are multiple options, take the one in common with the focal trial
    #all_trials$best_project[i] <- top_match[1]
    
    ## from the project name, get the id
    projId <- resp$content$results$counts$ranked_genotyping_projects[[1]]
    #all_trials$best_project_id[i] <- projId
    
    ## get the protocol name the same way
    req <- db$filter_geno_protocols(accessions = genos$germplasmDbId, verbose = FALSE)
    protId <- req$content$results$counts$ranked_genotyping_protocols[[1]]
    
    info <- wheat$vcf_archived_list(genotyping_protocol_id=protId, verbose = FALSE)
    if(nrow(info) > 1){
      info <- info |> #filter(project_id == projId, protocol_id == protId) |> 
        head(1) |> as.data.frame()
    }
    all_trials$best_protocol[i] <- info$protocol_name
    all_trials$best_protocol_id[i] <- info$protocol_id
    all_trials$best_project[i] <- info$project_name
    all_trials$best_project_id[i] <- info$project_id
    all_trials$vcf_file_name[i] <- info$file_name
    
    ## get the proportion of lines genotyped 
    nb_genos <- sum(resp$data$matches[,top_match[1]], na.rm=T)
    all_trials$prop_geno[i] <- 100*round(nb_genos / nrow(genos),1)
    # ## add the list of genotypes in the trial
    # all_trials$genos_list[i] <- list(genos$germplasmDbId)
  } else {
    all_trials$best_protocol[i] <- NA
  }
}
close(pb)
```

Export the table with genotyping platform info
```{r}
#TODO
```


# Explore output

Number of trials with genotyped lines for each program
```{r}
all_trials_geno <- all_trials |> 
  filter(!is.na(prop_geno) & prop_geno > 1) 
dim(all_trials_geno)
table(all_trials_geno$program)
```
Kansas doesn't have any trial with genotyped lines.

Check the genotyping platform diversity
```{r}
table(all_trials_geno$best_protocol)
intersect(manifests$protocol_name, unique(all_trials_geno$best_protocol))
```
# Download all vcf files

```{r}

for(prog in focal_prg[!focal_prg %in% "USDA-ARS Kansas"]){
  prog_trials <- all_trials_geno |> 
    filter(program == prog) |> 
    select(best_protocol_id, best_project_id, vcf_file_name) |> 
    distinct()
  write(paste0("Downloading ", nrow(prog_trials), " VCF files for program ", prog), stdout())
  tri <- unique(all_trials_focal |> filter(program == prog) |> pull(id))
  p2d <- here("data","genotypic",tri,"nonFocal")
  if(!dir.exists(p2d)) dir.create(p2d, recursive = TRUE)
  
  better_prog_name <- make_clean_names(prog,case="snake")
  for(i in 1:nrow(prog_trials)){
    wheat$vcf_archived(
      output =paste0(p2d,"/",better_prog_name,"_",i,".vcf.gz"),
      genotyping_protocol_id = prog_trials$best_protocol_id[i],
      genotyping_project_id =  prog_trials$best_project_id[i],
      file_name = prog_trials$vcf_file_name[i],
      verbose = FALSE
    )
  }
}
```
# Combine all vcf files into one per program

## Load vcf files

```{r}
#TODO
```


## Merge into one big file

```{r}
#TODO
```



## Format and impute
```{r}
#TODO
```

