---
title: "Pull genotype data from T3"
author: Charlotte Brault
output: html_document
date: '`r format(Sys.time(), "%d %B, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE)
```

# Libraries

```{r load libraries}
suppressPackageStartupMessages(library(here)) # root directory
suppressPackageStartupMessages(library(data.table)) # load data
suppressPackageStartupMessages(library(apercu)) # ap
suppressPackageStartupMessages(library(dplyr)) # formatting
suppressPackageStartupMessages(library(purrr)) # handle lists
## devtools::install_github("chabrault/HRSWGS") # custom package instal
#suppressPackageStartupMessages(library(HRSWGS)) # custom package
## devtools::install_github("TriticeaeToolbox/BrAPI.R")
suppressPackageStartupMessages(library(BrAPI)) # T3 connection
suppressPackageStartupMessages(library(DT)) # print table
suppressPackageStartupMessages(library(DT)) # print table
suppressPackageStartupMessages(library(vcfR)) # load vcf
```


# Set up root directory

```{r current dir}
here::i_am("analysis/pullGenoTrial.Rmd")
```


```{r}
trials_predictathon <- c("AWY1_DVPWA_2024","TCAP_2025_MANKS","25_Big6_SVREC_SVREC",
                         "OHRWW_2025_SPO","CornellMaster_2025_McGowan","24Crk_AY2-3",
                         "2025_AYT_Aurora","YT_Urb_25","STP1_2025_MCG")
```


# Create T3 connection

What's the difference between getBrAPIConnection and createBrAPIConnection?
```{r}
#wheat <- getBrAPIConnection("T3/Wheat")
db <- getBrAPIConnection("T3/Wheat")
wheat <- createBrAPIConnection("wheat.triticeaetoolbox.org", is_breedbase=TRUE)
#trials_df <- T3BrapiHelpers::getAllTrialMetaData(wheat,"Wheat")
```


# Get genotyping protocols that have data for any of the trial's accessions

## Extract genotype names from trials

From BrAPI reference page (https://triticeaetoolbox.github.io/BrAPI.R/#breedbase-functions)

```{r}
# Find accessions that were observed in these trials
geno_number <- map(trials_predictathon, \(x) {
  wheat$wizard("accessions", list(trials = x), verbose=F)$data$ids
}) 
names(geno_number) <- trials_predictathon
ap(geno_number)
sapply(geno_number,length)
```

We can also extract genotype names instead of numbers (replace `data$ids` by `data$names`)

## Get genotyping protocols

```{r}
# Get genotyping protocols that have data for any of the trial's accessions
geno_protocols <- map(geno_number, \(x)
                      wheat$wizard("genotyping_protocols", list(accessions = x), verbose=F)$data$ids)
names(geno_protocols) <- trials_predictathon
ap(geno_protocols)

geno_matches <- map(geno_number, \(x) db$filter_geno_protocols(x)$data$matches)
names(geno_matches) <- trials_predictathon
geno_matches_df <- bind_rows(geno_matches, .id="trial")
DT::datatable(geno_matches_df)
# geno_pro <- map(geno_number, \(x)
#   wheat$wizard("genotyping_projects", list(accessions = x), verbose=F)$data)
```
For the first trial (AWY1_DVPWA_2024), there is only one genotyping protocol (291 = ThermoFisher AgriSeq 5K), let's download a VCF for this one.

## Download vcf for one trial

```{r}
tri <- "AWY1_DVPWA_2024"
(prot <- geno_protocols[[tri]][1])
(geno_ids <- geno_number[[tri]])

proj <-  wheat$wizard("genotyping_projects", list(accessions = geno_ids), verbose=F)$data$ids

## check available files for this protocol
wheat$vcf_archived_list(genotyping_protocol_id=prot, verbose=T)

## define path to file to export
p2f <- here("data","genotypic",paste0(tri,"_genotypes_",prot,"protocol.vcf"))
## export vcf
wheat$vcf_archived(
  output=p2f,
  genotyping_protocol_id = prot,
  genotyping_project_id = proj,
  file_name = "2025-12-10_17:56:29_filevCTU",
  verbose = FALSE
)
```

There are two files for the same genotyping project and protocol => check if we need to combine them or if they are the same.

# Load downloaded vcf

```{r}
vcf <- list()
vcf.read <- vcfR::read.vcfR(file=p2f, verbose=FALSE)  
vcf$gt <- vcfR::extract.gt(vcf.read, element = "GT",
                           as.numeric=F, IDtoRowNames = T)
vcf$mrk.info <- tibble::as_tibble(vcfR::getFIX(vcf.read))
vcf$mrk.info <- vcf$mrk.info |>
  mutate(POS=as.numeric(POS)) |> arrange(CHROM, POS)
ap(vcf)
```


<!-- ```{r} -->
<!-- resp <- wheat$get("/programs", page="all") -->
<!-- programs <- sort(sapply(resp$combined_data, \(x) { x$programName })) -->

<!-- selected_breeding_program <- "University of Nebraska" -->
<!-- resp <- wheat$get("/studies", query=list(programName=selected_breeding_program), page="all") -->
<!-- trials <- lapply(resp$combined_data, \(x) { list(id=x$studyDbId, name=x$studyName) }) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- selected_trials <- c("","") -->
<!-- for ( trial_name in selected_trials ) { -->
<!--   resp <- wheat$get("/studies", query=list(studyName=trial_name)) -->
<!--   trial_metadata <- resp$data[[1]] -->
<!--   trial_id <- trial_metadata$studyDbId -->
<!--   location <- trial_metadata$locationName -->
<!--   planting_date <- trial_metadata$startDate -->
<!--   harvest_date <- trial_metadata$endDate -->
<!--   design <- trial_metadata$experimentalDesign$PUI -->
<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- geno <- wheat$wizard("genotyping_protocols", list(accessions = "LINKERT")) -->

<!-- #geno <- wheat$wizard("genotyping_protocols", list(accessions = accessions$data$ids)) -->
<!-- ``` -->

