---
title: "Choose potential training trials and archived VCFs"
author: "Jean-Luc Jannink"
date: "2026-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(lubridate)
require(stringr)
require(janitor)
require(readr)
require(BrAPI)
require(T3BrapiHelpers)
require(here)
here::i_am("analysis/T3Predictathon2026_Choose_Potential_Training_Trials_and_Archived_VCFs.Rmd")
source(
  here::here("code",
             "T3Predictathon2026_Functions_to_choose_trials_and_protocols.R"))
source(
  here::here("code",
             "T3Predictathon2026_Functions_to_work_with_VCFs.R"))

wheat_conn <- BrAPI::getBrAPIConnection("T3/Wheat")
```

#### Start with metadata from all the phenotyping trials
NOTE: recreate this metadata file on March 20th or so.  

```{r Check for metadata}
# Check if the metadata file is there. If not, create it

if (!file.exists(here::here("data", "All_wheat_trial_metadata.csv"))){
  trials_metadata <- T3BrapiHelpers::getAllTrialMetaData(wheat_conn, "wheat")
  readr::write_csv(trials_metadata,
                   file=here::here("data", "All_wheat_trial_metadata.csv"))
}
```

#### Test genotyping project
Note: choose_training_trials finds trials based on the accessions listed in
those trials.  If data for the trial was never uploaded, the trial will be 
returned even though the trial is not useful.  

You can use this same function for other studies besides the Predictathon trials
that is, trials that _have_ observations so that you can test whether this
approach gives good results.  
```{r Find and save genotyping project for the test trial}
choose_training_trials <- function(study_id,
                                   min_primary_germ_common=3,
                                   min_secondary_germ_common=12){
  cat("Finding reasonable training trials for focal trial", study_id, "\n")
  cat("min_primary_germ_common", min_primary_germ_common, "   ")
  cat("min_secondary_germ_common", min_secondary_germ_common, "\n")

  # Wrapper for wizard to find all trials where an accession was evaluated
  get_trial_ids_from_germplasm_id <- function(germplasm_id, brapi_conn){
    return(brapi_conn$wizard(data_type = "trials",
                             filters=list(accessions=germplasm_id))$data$ids)
  }

  # Open the connection to T3/Wheat
  wheat_conn <- BrAPI::getBrAPIConnection("T3/Wheat")
  
  # Get the accessions evaluated in the focal trial
  focal_germplasm <- wheat_conn$wizard(data_type = "accessions",
                                       filters=list(trials=study_id))$data$ids
  
  # For each focal accession, what trials was it evaluated in?
  primary_studies <- focal_germplasm |>
    purrr::map(.f=get_trial_ids_from_germplasm_id,
               brapi_conn=wheat_conn,
               .progress=TRUE)
  
  # Find trials that evaluated more than `min_primary_germ_common` accessions
  # These should be trials that evaluate similar germplasm to the focal trial
  # This is the primary pool of trials to use in training
  primary_studies_tabyl <- janitor::tabyl(unlist(primary_studies)) |>
    dplyr::select(-percent) |>
    dplyr::rename(primary_study_db_id=`unlist(primary_studies)`) |>
    dplyr::filter(primary_study_db_id != study_id) |>
    dplyr::filter(n >= min_primary_germ_common)
  
  primary_study_db_ids <- unlist(primary_studies_tabyl$primary_study_db_id)
  cat("length primary_study_db_ids", length(primary_study_db_ids), "\n")
  
  # Maybe no other trials evaluate the same accessions. Then, return NA.
  if (length(primary_study_db_ids) == 0){
    return(list(study_id=study_id,
                min_primary_germ_common=min_primary_germ_common,
                min_secondary_germ_common=min_secondary_germ_common,
                focal_germplasm=focal_germplasm,
                primary_study_db_ids=NA,
                primary_germplasm=NA,
                secondary_study_db_ids=NA,
                secondary_germplasm=NA))
  }
  
  # What accessions were evaluated in this primary pool
  primary_germplasm <- wheat_conn$wizard(
    data_type = "accessions",
    filters=list(trials=primary_study_db_ids))$data$ids
  
  # In what trials were these primary training accessions evaluated
  secondary_studies <- primary_germplasm |>
    purrr::map(.f=get_trial_ids_from_germplasm_id,
               brapi_conn=wheat_conn,
               .progress=TRUE)
  
  # Find trials that evaluated more than `min_secondary_germ_common` accessions
  # These should be trials that evaluate similar germplasm to the primary trials
  # This is the secondary pool of trials to use in training
  secondary_studies_tabyl <- janitor::tabyl(unlist(secondary_studies)) |>
    dplyr::select(-percent) |>
    dplyr::rename(secondary_study_db_id=`unlist(secondary_studies)`) |>
    dplyr::filter(secondary_study_db_id != study_id) |>
    dplyr::filter(n >= min_secondary_germ_common)
  
  # Need to use setdiff here because all of the primary trials in here too
  secondary_study_db_ids <- setdiff(
    unlist(secondary_studies_tabyl$secondary_study_db_id),
    primary_study_db_ids)
  cat("length secondary_study_db_ids", length(secondary_study_db_ids), "\n")

  # Maybe no other trials evaluate the same accessions. Then, return NA.
  if (length(secondary_study_db_ids) == 0){
    return(list(study_id=study_id,
                min_primary_germ_common=min_primary_germ_common,
                min_secondary_germ_common=min_secondary_germ_common,
                focal_germplasm=focal_germplasm,
                primary_study_db_ids=primary_study_db_ids,
                primary_germplasm=primary_germplasm,
                secondary_study_db_ids=NA,
                secondary_germplasm=NA))
  }
  
  # What accessions were evaluated in this secondary pool
  secondary_germplasm <- setdiff(
    wheat_conn$wizard(data_type = "accessions",
                      filters=list(trials=secondary_study_db_ids))$data$ids,
    primary_germplasm)
  
  # DONE!
  return(list(study_id=study_id,
              min_primary_germ_common=min_primary_germ_common,
              min_secondary_germ_common=min_secondary_germ_common,
              focal_germplasm=focal_germplasm,
              primary_study_db_ids=primary_study_db_ids,
              primary_germplasm=primary_germplasm,
              secondary_study_db_ids=secondary_study_db_ids,
              secondary_germplasm=secondary_germplasm))
}

trial_names_in_order <- c("2025_AYT_Aurora", "24Crk_AY2-3",
                          "25_Big6_SVREC_SVREC", "CornellMaster_2025_McGowan",
                          "YT_Urb_25", "AWY1_DVPWA_2024", "OHRWW_2025_SPO",
                          "TCAP_2025_MANKS", "STP1_2025_MCG")

# These are the studyDbIds of the Predictathon trials
predict_trial_vec_Wheat <- c("10673", "10674", "10675", "10676",
                             "10677", "10678", "10679", "10680", "10681")

# Different numbers of accessions in common for primary and secondary trials
# for the different Predictathon trials just based on what's available
parameters_to_find_related_trials <- tibble::tibble(
  study_id=predict_trial_vec_Wheat,
  min_primary_germ_common=c(3, 3, 4, 3, 3, 6, 6, 4, 3),
  min_secondary_germ_common=c(12, 4, 15, 12, 15, 20, 20, 20, 6))

# It's nice not to use purrr here because it's slow.  I takes ~ 30 minutes per
# Predictathon trial 
training_trial_info <- list()

for (i in 1:nrow(parameters_to_find_related_trials)){
  cat(unlist(parameters_to_find_related_trials[i,]), "\n")
  training_trial_info[[i]] <- choose_training_trials(
      study_id=parameters_to_find_related_trials$study_id[i],
      min_primary_germ_common=
        parameters_to_find_related_trials$min_primary_germ_common[i],
      min_secondary_germ_common=
        parameters_to_find_related_trials$min_secondary_germ_common[i])

  saveRDS(training_trial_info, file=here::here("output", "Training_Trial_Info.rds"))
}

# Figure out the numbers of trials and accessions in training trial inro
n_trials_accessions <- sapply(training_trial_info, function(lv){sapply(lv, length)})

phenotype_file_names <- purrr::map(training_trial_info, 
                                   purrr::safely(download_training_trials),
                                   .progress=TRUE)

genotype_file_names <- purrr::map(training_trial_info,
                                  purrr::safely(download_useful_vcfs),
                                  .progress=TRUE)
```

#### For trials that were _not_ downloaded
Make sure that that happened because they have no yield information as opposed
to some problem with the database or BrAPI or whatever.  

```{r Double check that trials not in the folder do not have observations}
files_missing_in_folder <- function(i){
  study_id <- training_trial_info[[i]]$study_id
  folder_path <- here::here("data", paste0("study", study_id))
  files <- list.files(folder_path)
  file_study_ids <- gsub("primary_phenotypes_", "", files)
  file_study_ids <- gsub("secondary_phenotypes_", "", file_study_ids)
  file_study_ids <- gsub(".csv", "", file_study_ids)
  training_trials <- union(training_trial_info[[i]]$primary_study_db_ids,
                           training_trial_info[[i]]$secondary_study_db_ids) |> 
    as.character()
  return(setdiff(training_trials, file_study_ids))
}

files_missing <- lapply(1:9, files_missing_in_folder)

# This function just looks to see if a trial has yield observations
check_missing_files <- function(study_id){
  yield_phenotypes <- wheat_conn$search(
    "observations", 
    body=list(studyDbIds=c(study_id),
              observationVariableDbIds=c(84527)))$combined_data |>
    purrr::map(function(sl){ sl$season <- sl$season[[1]]; sl}) |>
    dplyr::bind_rows()
  return(nrow(yield_phenotypes))
}

for (i in 1:9){
  tst <- purrr::map(files_missing[[i]],
                    purrr::safely(check_missing_files),
                    .progress=TRUE)
  tst <- tst |> sapply(function(sl) return(sl$result)) |> unlist()
  
  # This number should match the difference between the numbers of trials
  # in n_trials_accessions and the number of trials that were downloaded
  print(sum(tst == 0))
  print(which(tst > 0))
}
```

#### The observations from the trials were saved to a data folder
But we still need the trial metadata

```{r Get and save training trial metadata}
training_trial_metadata <- list()
for (study_id in predict_trial_vec_Wheat){
  cat(study_id, "\n")
  folder_path <- here::here("data", paste0("study", study_id))
  files <- list.files(folder_path)
  file_study_ids <- gsub("primary_phenotypes_", "", files)
  file_study_ids <- gsub("secondary_phenotypes_", "", file_study_ids)
  file_study_ids <- gsub(".csv", "", file_study_ids)
  
  file_study_ids <- file_study_ids[!is.na(file_study_ids)]
  trial_metadata <- T3BrapiHelpers::getTrialMetaDataFromTrialVec(file_study_ids,
                                                                 wheat_conn)
  trial_lat_long <- T3BrapiHelpers::getLatLongElevFromLocationVec(
    trial_metadata$location_name, wheat_conn, id_or_name="name")
  trial_metadata <- merge(trial_metadata, trial_lat_long) |> 
    dplyr::select(study_name, study_db_id, program_name, study_type, 
                  location_name, location_db_id, 
                  latitude, longitude, elevation,
                  start_date, end_date, experimental_design)
  
  training_trial_metadata[[which(predict_trial_vec_Wheat == study_id)]] <- trial_metadata
  
  # Make sure the subdirectory is there
  subdir <- here::here("data", paste0("study", study_id))
  if (!dir.exists(subdir)) dir.create(subdir)

  # This prefix targets the download to the specific folder
  file_name <- here::here(
    "data",
    paste0("study", study_id, "/training_trial_metadata_for_", study_id, ".csv"))
  
  readr::write_csv(trial_metadata, file = file_name)
}
```
